---
title: "Salary_Data"
author: "Sanjyot Godbole"
date: "February 12, 2019"
output: 
  html_document:
    toc: true
    theme: united
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Clearing the environment
```{r ClearingEnvironment}
rm(list = ls(all=TRUE))
```

##Loading the dataset 
```{r}

Train_Data = read.csv("Train-1550224040162.csv", header = T,sep = ",")
dataToPredict = read.csv("Test-1550224068068.csv", header = T,sep = ",")
dim(Train_Data)
dim(dataToPredict)

```

##Descriptive Analysis of Train_Data
```{r}
str(Train_Data)
head(Train_Data)
tail(Train_Data)
summary(Train_Data)
names(Train_Data)
```

## Pre-processing 
### Changing the datatypes of some columns

```{r }
Train_Data$CollegeTier = as.factor(Train_Data$CollegeTier)
Train_Data$CollegeCode = as.factor(Train_Data$CollegeCode)
Train_Data$CityCode = as.factor(Train_Data$CityCode)
Train_Data$CityTier = as.factor(Train_Data$CityTier)


Train_Data$Year.Of.Twelth.Completion = as.numeric(Train_Data$Year.Of.Twelth.Completion)
Train_Data$Year.of.Graduation.Completion = as.numeric(Train_Data$Year.of.Graduation.Completion)
Train_Data$Score.in.Logical.skill = as.numeric(Train_Data$Score.in.Logical.skill)
Train_Data$Score.in.Quantitative.ability = as.numeric(Train_Data$Score.in.Quantitative.ability)
Train_Data$Score.in.Quantitative.ability = as.numeric(Train_Data$Score.in.Quantitative.ability)
Train_Data$Score.in.ComputerProgramming = as.numeric(Train_Data$Score.in.ComputerProgramming)
Train_Data$Score.in.ElectronicsAndSemicon = as.numeric(Train_Data$Score.in.ElectronicsAndSemicon)
Train_Data$Score.in.ComputerScience = as.numeric(Train_Data$Score.in.ComputerScience)
Train_Data$Score.in.MechanicalEngg = as.numeric(Train_Data$Score.in.MechanicalEngg)
Train_Data$Score.in.ElectricalEngg = as.numeric(Train_Data$Score.in.ElectricalEngg)
Train_Data$Score.in.TelecomEngg = as.numeric(Train_Data$Score.in.TelecomEngg)
Train_Data$Score.in.CivilEngg = as.numeric(Train_Data$Score.in.CivilEngg)


Train_Data$Date.Of.Birth = as.Date(Train_Data$Date.Of.Birth)

```



## Repeating all the steps on data to predict
```{r}
dataToPredict$CollegeTier = as.factor(dataToPredict$CollegeTier)
dataToPredict$CollegeCode = as.factor(dataToPredict$CollegeCode)
dataToPredict$CityCode = as.factor(dataToPredict$CityCode)
dataToPredict$CityTier = as.factor(dataToPredict$CityTier)


dataToPredict$Year.Of.Twelth.Completion = as.numeric(dataToPredict$Year.Of.Twelth.Completion)
dataToPredict$Year.of.Graduation.Completion = as.numeric(dataToPredict$Year.of.Graduation.Completion)
dataToPredict$Score.in.Logical.skill = as.numeric(dataToPredict$Score.in.Logical.skill)
dataToPredict$Score.in.Quantitative.ability = as.numeric(dataToPredict$Score.in.Quantitative.ability)
dataToPredict$Score.in.Quantitative.ability = as.numeric(dataToPredict$Score.in.Quantitative.ability)
dataToPredict$Score.in.ComputerProgramming = as.numeric(dataToPredict$Score.in.ComputerProgramming)
dataToPredict$Score.in.ElectronicsAndSemicon = as.numeric(dataToPredict$Score.in.ElectronicsAndSemicon)
dataToPredict$Score.in.ComputerScience = as.numeric(dataToPredict$Score.in.ComputerScience)
dataToPredict$Score.in.MechanicalEngg = as.numeric(dataToPredict$Score.in.MechanicalEngg)
dataToPredict$Score.in.ElectricalEngg = as.numeric(dataToPredict$Score.in.ElectricalEngg)
dataToPredict$Score.in.TelecomEngg = as.numeric(dataToPredict$Score.in.TelecomEngg)
dataToPredict$Score.in.CivilEngg = as.numeric(dataToPredict$Score.in.CivilEngg)


dataToPredict$Date.Of.Birth = as.Date(dataToPredict$Date.Of.Birth)
```

##Handling missing values
### Train Dataset
**Using central Imputation**
```{r}
#Checking dimensions before handling missing values
dim(Train_Data)

#Checking total number of missing values in the dataset
sum(is.na(Train_Data))

#Checking feature-wise missing values
for (feature in names(Train_Data)) {
    missing <- sum(is.na(Train_Data[,feature]))
    if (missing > 0) {
        print(c(feature,missing))
      
    }
}

# All the 18 na values are present in Date.Of.Birth column

#Counting the remaining missing values
library(dplyr)
nrow(filter(Train_Data, School.Board.in.Tenth == "0"))
nrow(filter(Train_Data, Board.in.Twelth == "0"))
nrow(filter(Train_Data, Score.in.Domain == "-1"))


# There as 11036 missing values

#Assing NA values 

Train_Data$School.Board.in.Tenth[Train_Data$School.Board.in.Tenth=="0"] <- NA
Train_Data$Board.in.Twelth[Train_Data$Board.in.Twelth=="0"] <- NA
Train_Data$Score.in.Domain[Train_Data$Score.in.Domain=="-1"] <- NA

##########################################################################################

#Imputting remanining values using centralimputation
library(DMwR)

Train_Data_no_NA<-centralImputation(Train_Data)

#Train_Data_no_NA<-knnImputation(Train_Data)
#View(Train_Data_no_NA)

#Checking the number of missing values left after KNN imputation
sum(is.na(Train_Data_no_NA))
dim(Train_Data_no_NA)

#########################################################################################

#Saving a CSV file of imputed data to avoid imputation time while re-running the code  
write.csv(Train_Data_no_NA, "Train_Data_no_NA_new.csv", row.names=F)

```

### dataToPredict Dataset
**Using Central Imputation**
```{r}
#Checking dimensions before handling missing values
dim(dataToPredict)

#Checking total number of missing values in the dataset
sum(is.na(dataToPredict))

#Checking feature-wise missing values
for (feature in names(dataToPredict)) {
    missing <- sum(is.na(dataToPredict[,feature]))
    if (missing > 0) {
        print(c(feature,missing))
      
    }
}

# All the 12 na values are present in Date.Of.Birth column

#Counting the remaining missing values
library(dplyr)
nrow(filter(dataToPredict, School.Board.in.Tenth == "0"))
nrow(filter(dataToPredict, Board.in.Twelth == "0"))
nrow(filter(dataToPredict, Score.in.Domain == "-1"))


# There as 4575 missing values

#Assing NA values 

dataToPredict$School.Board.in.Tenth[dataToPredict$School.Board.in.Tenth=="0"] <- NA
dataToPredict$Board.in.Twelth[dataToPredict$Board.in.Twelth=="0"] <- NA
dataToPredict$Score.in.Domain[dataToPredict$Score.in.Domain=="-1"] <- NA

##########################################################################################

#Imputting remanining values using centralimputation
library(DMwR)

dataToPredict_no_NA<-centralImputation(dataToPredict)

#dataToPredict_no_NA<-knnImputation(dataToPredict)
#View(dataToPredict_no_NA)

#Checking the number of missing values left after KNN imputation
sum(is.na(dataToPredict_no_NA))
dim(dataToPredict_no_NA)

#########################################################################################

#Saving a CSV file of imputed data to avoid imputation time while re-running the code  
write.csv(dataToPredict_no_NA, "dataToPredict_no_NA_new.csv", row.names=F)

```

## Scaling the data
```{r}
library(vegan)
Train_Data_no_NA_num = select_if(Train_Data_no_NA, is.numeric)
Train_Data_no_NA_num$Candidate.ID = NULL
Train_Data_no_NA_num$Pay_in_INR = NULL

dataToPredict_no_NA_num = select_if(dataToPredict_no_NA, is.numeric)
dataToPredict_no_NA_num$Candidate.ID = NULL

Train_Data_no_NA_num_scaled =decostand(select_if(Train_Data_no_NA_num, is.numeric), method = "range")
dataToPredict_no_NA_num_scaled =decostand(select_if(dataToPredict_no_NA_num, is.numeric), method = "range")

#Merging the remaining data
Train_Data_no_NA_not_num = Train_Data_no_NA %>% select_if(funs(!is.numeric(.)))
Train_Data_no_NA_num_scaled = cbind(Train_Data_no_NA$Candidate.ID, Train_Data_no_NA$Pay_in_INR,Train_Data_no_NA_not_num, Train_Data_no_NA_num_scaled)

dataToPredict_no_NA_not_num = dataToPredict_no_NA %>% select_if(funs(!is.numeric(.)))
dataToPredict_no_NA_num_scaled = cbind(dataToPredict_no_NA$Candidate.ID,dataToPredict_no_NA_not_num, dataToPredict_no_NA_num_scaled)

names(Train_Data_no_NA_num_scaled)[names(Train_Data_no_NA_num_scaled) == 'Train_Data_no_NA$Candidate.ID'] <- 'Candidate.ID'

names(Train_Data_no_NA_num_scaled)[names(Train_Data_no_NA_num_scaled) == 'Train_Data_no_NA$Pay_in_INR'] <- 'Pay_in_INR'

names(dataToPredict_no_NA_num_scaled)[names(dataToPredict_no_NA_num_scaled) == 'dataToPredict_no_NA$Candidate.ID'] <- 'Candidate.ID'

#Saving a CSV file of imputed data to avoid imputation time while re-running the code 

write.csv(Train_Data_no_NA_num_scaled, "Train_Data_no_NA_num_scaled_V2.csv", row.names=F)

write.csv(dataToPredict_no_NA_num_scaled, "dataToPredict_no_NA_num_scaled_V2.csv", row.names=F)

```


## Exploratory Data Analysis - 

###Checking the distribution of "Pay_in_INR"
```{r}
library(ggplot2)

ggplot(Train_Data_no_NA, aes(Pay_in_INR)) + #, fill= as.factor(target) )) +
  theme_bw() +
  #facet_wrap(gender ~ qualification) +
  #facet_grid(gender ~ qualification) +
  geom_histogram(binwidth = 100000) +
  labs(y = "individual count",
       x = "Pay_in_INR(binwidth = 100000)",
       title = "Pay_in_INR distribution")

```

###Checking the skewness of all the numerical data for checking possible transformations
```{r}
library(tidyr)
library(purrr)

Train_Data_no_NA_num_scaled %>%
  purrr::keep(is.numeric) %>%
  tidyr::gather() %>%
  ggplot2::ggplot(aes(value))+
  facet_wrap(~key, scales='free')+
  geom_histogram()


```
### Checking correlation between all the attributes
```{r}
 
cor_data = cor(select_if(Train_Data_no_NA_num_scaled, is.numeric))

library(corrplot)
corrplot(cor_data, method = "number", tl.cex = 0.70)

```


###Scatterplot to check gender wise corelation between "Year.Of.Twelth.Completion" and "Year.of.Graduation.Completion"
```{r }


library(ggplot2)
ggplot(Train_Data_no_NA_num_scaled, aes(Year.Of.Twelth.Completion, Year.of.Graduation.Completion, color = Gender)) +
geom_point(shape = 16, size = 6, show.legend = TRUE) 
```
